---
title: "Statistical Modeling Lecture"
author: "Devin J. Cornell"
date: "10/19/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(modelr) # this one is new!
library(lubridate)
```

# Our Dataset: Twitter data from the last Lab!

```{r}
load(url('https://dssoc.github.io/datasets/congress.RData'))
load(url('https://dssoc.github.io/datasets/senator_tweets.RData'))
senator_tweet_sample %>% head()
```

# Advantages to Visualization

## Start with basic Scatter Plot

```{r}
# basic scatter plot
senator_tweet_sample %>% 
  ggplot(aes(x=favorite_count, y=retweet_count)) +
    geom_point()
```

## Can we improve it? Try adjusting alpha

```{r}
# basic scatter plot
senator_tweet_sample %>% 
  ggplot(aes(x=favorite_count, y=retweet_count)) +
    geom_point(aes(alpha=0.2))
```



## Add trendlines


```{r}
# basic scatter plot
senator_tweet_sample %>% 
  ggplot(aes(x=favorite_count, y=retweet_count)) +
    geom_point() +
    geom_smooth(method='lm', formula= y ~ x)
```


## Try log version for better visualization


```{r}
# easier to see if we take the log
senator_tweet_sample %>% 
  ggplot(aes(x=log(favorite_count), y=log(retweet_count))) +
    geom_point()
```



## Add trendline to lot plot

```{r}
senator_tweet_sample %>% 
  ggplot(aes(x=log(favorite_count), y=log(retweet_count))) +
    geom_point(aes(alpha=0.1)) + 
    geom_smooth(method='lm', formula= y ~ x)
```


## What is off about this plot?

### HINT: why do so many tweets have zero favorites?

```{r}
senator_tweet_sample %>% 
  group_by(is_retweet) %>% 
  summarize(mean(favorite_count))
```


## Now improve the trendline information with that knowledge

```{r}
senator_tweet_sample %>% 
  filter(!is_retweet) %>% 
  ggplot(aes(x=favorite_count, y=retweet_count)) +
    geom_point(aes(alpha=0.1)) + 
    geom_smooth(method='lm', formula= y ~ x)
```


## How do we get quantitative information about the relationship between these two variables?


### try a correlation first.

```{r}
tmp <- senator_tweet_sample %>% 
  filter(!is_retweet)

cor.test(tmp$retweet_count, tmp$favorite_count)
```


### now try creating a linear model with `lm`

```{r}
tmp <- senator_tweet_sample %>% 
  filter(!is_retweet)

m <- lm(retweet_count ~ favorite_count, data=tmp)
m
m %>% summary()
```


## How can we visualize models?


### One approach: create predictions from your actual dataset

```{r}
tmp$pred <- m %>% predict(tmp)
tmp %>% ggplot() +
  geom_line(aes(x=favorite_count, y=pred)) +
  geom_point(aes(x=favorite_count, y=retweet_count, alpha=0.1))

# check it out - R^2 is just based on the differences between predictions and observations
a <- (tmp$retweet_count - tmp$pred)^2 %>% sum()
b <- (tmp$retweet_count - mean(tmp$retweet_count))^2 %>% sum()
rsq <- 1-a/b

# this is the same as we see above in the linear model summary
rsq
```

### Another approach: create a synthetic dataset from which to generate predictions

```{r}
pred_df <- data.frame(favorite_count=1:5000)
pred_df$pred <- m %>% predict(pred_df)
pred_df %>% ggplot(aes(x=favorite_count, y=pred)) +
  geom_line()
```


# Which factors determine the number of Retweets a given Tweet receives?

## Try variables in the Tweet dataset

```{r}
tmp <- senator_tweet_sample %>% 
  mutate(is_weekend=wday(created_at, label = TRUE) %in% c('Sun', 'Sat'))
tmp %>% head()

m <- lm(retweet_count ~ is_weekend + is_quote, data=tmp)
m %>% summary()
```


## Now augment with demographic characteristics from the `congress` dataframe

```{r}
# get info about congress members
cdf <- congress %>% 
  left_join(congress_contact, by='bioguide_id') %>% 
  mutate(twitter=tolower(twitter))

model_df <- senator_tweet_sample %>% 
  mutate(is_weekend=wday(created_at, label = TRUE) %in% c('Sun', 'Sat')) %>% 
  mutate(screen_name=tolower(screen_name)) %>% 
  left_join(cdf, by=c('screen_name'='twitter')) %>% 
  drop_na(gender) %>% 
  mutate(is_repub=party=='Republican', age=2022-birthyear)
  
model_df %>% head()
lm(favorite_count ~ is_weekend + is_quote + gender + age + is_repub, data=model_df) %>% summary()
```

## does the effect of age on number of favorites vary by gender?

```{r}
# try an interaction between gender and age
m_interact <- lm(favorite_count ~ is_weekend + is_quote + is_repub + gender*age, data=model_df)
m_interact %>% summary()
```

## how can we visualize this model?

```{r}
# make a synthetic dataset from which to create predictions
pred_df <- model_df %>% 
  data_grid(gender, age) %>% 
  mutate(is_weekend=F, is_quote=F, is_repub=F)
pred_df$pred <- m_interact %>% predict(pred_df)

pred_df %>% ggplot(aes(x=age, y=pred, color=gender)) +
  geom_line() +
  ylab('Predicted Favorite Count') +
  xlab('Author Age')

```












