---
title: "Lab #5: Basics of Social Network Analysis"
author: "Data Science and Society (Sociology 367)"
date: ""
output: html_document
---


In this lab, we will practice the fundamentals of social network analysis. Edit the Lab #5 markdown file ([link to file](/assignments/Lab_5.Rmd)) for your submission. Remember to change the "author" field above to your name. **Send the markdown file (.Rmd) to your TA via direct message on Slack.** Be sure to do the required reading!

**Required readings:** 

* [Intro to Network Analysis with R, by Jesse Sadle](https://www.jessesadler.com/post/network-analysis-with-r/). 
* [Network analysis with R and igraph: NetSci X Tutorial (Parts 2-7), by 
Katya Ognyanova](https://kateto.net/networks-r-igraph)

[R for Data Science: Basic Programming (17-21)](https://r4ds.had.co.nz/program-intro.html).

**Optional reading:** 

* [Official documentation for `igraph` package](https://igraph.org/r/doc/)
* [Official documentation for `ggraph` package](https://cran.r-project.org/web/packages/ggraph/ggraph.pdf)
* [Introduction to `ggraph` layouts](https://www.data-imaginist.com/2017/ggraph-introduction-layouts/)


### Required Datasets

The network data that you will be drawing on for this lab can be downloaded from the course website: 

(1) `canidates_flw_el`: [dnc_canidates_flw_el.Rdata](/datasets/dnc_candidates_flw_el.Rdata) is an edgelist of the top five DNC cadidates for president (according to delegates won as of February 15, 2020), and the accounts they follow on Twitter.

(2) `canidates_flw_am`: [dnc_canidates_flw_am.Rdata](/datasets/dnc_candidates_flw_am.Rdata) is an adjacency matrix of the top five five DNC cadidates for president (according to delegates won as of February 15, 2020), where ties/edges are the number of accounts they mutually follow on Twitter.  


**Load the datasets:**
```{r message=FALSE}
#library(igraph)
#library(ggraph)
#library(tidyverse)
#load('../datasets/dnc_candidates_flw_am.Rdata')
#load('../datasets/dnc_candidates_flw_el.Rdata')
library(igraph)
library(ggraph)
library(tidyverse)
load('../datasets/dnc_candidates_flw_am.Rdata')
load('../datasets/dnc_candidates_flw_el.Rdata')
```
<br/><br/>

### Exercises
<br/>

**1. Describe what nodes/vertices and links/edges are. What kinds of things in the world might we attempt to model as networks?**

```
# Yor answer here or in text.
```
<br/>

**2. What kinds of questions would you be able to answer with networks that you might not be able to answer using analysis techniques we have covered so far?**
```
# Yor answer here or in text.
```
<br/>

**3. Describe the data in the `candidates_flw_am` adjacency matrix and the `candidates_flw_el` edgelist. How do the two networks differ? Do they represent directed or undirected edges? Do the edges have weights?**
```
# Yor answer here or in text.
```
<br/>

**4. Describe three different metrics (e.g. betweenness centrality, degree, etc) that can be used to summarize networks or specific nodes/vertices within a network.**
```{r}
# Your answer here
```
<br/>


**5. Using the `candidates_flw_am` *adjacency matrix*, create a graph (network) object. Make sure to include information about the edge weights (i.e. the number of accounts shared in common). Hints: (1) you may need to use the `as.matrix` function to the dataframe before converting to igraph if you are having problems with data types; (2) you may need to look through `igraph`'s official documentation (see link under "optional reading" above) to decide which function will convert adjacency matrices to igraph objects.**
```{r}
# Yor answer here or in text.
g <- graph_from_adjacency_matrix(adjmatrix=as.matrix(candidates_flw_am), weighted=TRUE, mode='undirected')
g
edge_attr(g)
```
<br/>


**6. Using the graph object that you created, indicate which candidate has the highest betweenness centrality, taking into account edge weight. If you had to write this finding in a research paper, how would you describe its meaning to a social science audience?**
```{r}
# Yor answer here or in text.
betweenness(g)
```
<br/>


**7. Make a visual of your graph object using `ggraph`, with edges scaled by weight (i.e. the number of accounts that the two candidates mutually follow). If you decide to make this graph in Gephi, feel free to submit a separate image of your final visual in addition to your .Rmd file.**
```{r}
# Yor answer here or in text.
E(g)$weights
ggraph(g, layout='kk') + 
  geom_node_point() + 
  geom_edge_link(aes(width=weight, alpha=0.05)) +
  #geom_edge_density(aes(fill=weight)) + 
  geom_node_text(aes(label=name))
```
<br/>

**8. Now we turn our attention to two-mode networks. Create an igraph object from the `candidates_flw_el` edgelist, and calculate the degree ONLY for each candidate (not their followers).**
```{r}
# Yor answer here or in text.
g2 <- graph_from_data_frame(candidates_flw_el)
degree(g2, v=unique(candidates_flw_el$candidate))
```
<br/>


**9. In last week's lab excerices, you were asked to indicate possible research topics that you would be interested in pursuing for your final project. Choose one of your topics and expand on it in more detail (i.e. at least one paragraph).**
```{r}
# your answer here
```



