---
title: "Lab #5: Basics of Social Network Analysis"
author: "Data Science and Society (Sociology 367)"
date: ""
output: html_document
---


In this lab, we will practice the fundamentals of social network analysis. Edit the Lab #5 markdown file ([link to file](https://github.com/dssoc/dssoc.github.io/raw/master/assignments/Lab_5.Rmd)) for your submission. Remember to change the "author" field above to your name. **Send the markdown file (.Rmd) to your TA via direct message on Slack.** Be sure to do the required reading!

**Required reading:** [R for Data Science: Basic Programming (17-21)](https://r4ds.had.co.nz/program-intro.html).

**Optional reading:** [Intro to Network Analysis with R, by Jesse Sadle](https://www.jessesadler.com/post/network-analysis-with-r/).


### Required Datasets

The network data that you will be drawing on for this lab can be downloaded from the course website: 

(1) `canidates_flw_el`: [dnc_canidates_flw_el.Rdata](https://github.com/dssoc/dssoc.github.io/raw/master/datasets/dnc_candidates_flw_el.Rdata) is an edgelist of the top five DNC cadidates for president (according to delegates won as of February 15, 2020), and the accounts they follow on Twitter.

(2) `canidates_flw_am`: [dnc_canidates_flw_am.Rdata](https://github.com/dssoc/dssoc.github.io/raw/master/datasets/dnc_candidates_flw_am.Rdata) is an adjacency matrix of the top five five DNC cadidates for president (according to delegates won as of February 15, 2020), where ties/edges are the number of accounts they mutually follow on Twitter.  


**Load the datasets:**
```{r}
#library(igraph)
#library(ggraph)
#library(tidyverse)
#load('../datasets/dnc_candidates_flw_am.Rdata')
#load('../datasets/dnc_candidates_flw_el.Rdata')
```
<br/><br/>

### Exercises
<br/>

**1. In class you learned that there are several ways to store network data. Describe the difference between an edgelist and an adjacency matrix.**

```{r}
# Yor answer here or in text.
```
<br/>

**2. You also learned about one- and two-mode networks. Describe the difference.**
```{r}
# Yor answer here or in text.
```
<br/>

**3. Does the `candidates_flw_am` adjacency matrix contain one- or two-mode network information? What about the `candidates_flw_el` edgelist?**
```{r}
# Yor answer here or in text.
```
<br/>

**4. Describe three different metrics (e.g. betweenness centrality, degree, etc) that can be used to summarize networks or specific nodes/vertices within a network.**
```{r}
# Your answer here
```
<br/>


**5. Using the `candidates_flw_am` adjacency matrix (a data frame), create a graph (network) object. Make sure to include information about the edge weights (i.e. the number of accounts shared in common). Note: you may need to use the `as.matrix` function to the dataframe first if you are having data type problems.**
```{r}
# Yor answer here or in text.
```
<br/>


**6. Using the graph object that you created, indicate which candidate has the highest betweenness centrality, taking into account edge weight. If you had to write this finding in a research paper, how would you describe its meaning to a social science audience?**
```{r}
# Yor answer here or in text.
```
<br/>


**7. Make a visual of your graph object using `ggraph`, with edges scaled by weight (i.e. the number of accounts that the two candidates mutually follow). If you decide to make this graph in Gephi, feel free to submit a separate image of your final visual in addition to your .Rmd file.**
```{r}
# Yor answer here or in text.
```
<br/>

**8. Now we turn our attention to two-mode networks. Create an igraph object from the `candidates_flw_el` edgelist, and calculate the degree for each candidate.**
```{r}
# Yor answer here or in text.
```
<br/>


**9. In last week's lab excerices, you were asked to indicate possible research topics that you would be interested in pursuing for your final project. Choose one of your topics and expand on it in more detail (i.e. at least one paragraph).**
```{r}
# your answer here
```



