---
---
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab 6: Working with APIs

[**Link to updated lab**](https://dssoc.github.io/assignments/Lab_6_updated.Rmd)

[**Lab Instructions**](https://dssoc.github.io/assignments/lab_instructions.html)

In this lab we will be learning about APIs.

See the "Instructions" section of the [Introduction to Lab Assignments](https://dssoc.github.io/assignments/lab_instructions.html) page for more information about the labs. That page also gives descriptions for the datasets we will be using.

**Required reading:**

**Optional resources:**

-   [Applying for a Twitter developer account](https://www.youtube.com/watch?v=LTC8laBJGP0&t=967s)
-   [rtweet Package Documentation](https://cran.r-project.org/web/packages/rtweet/rtweet.pdf)
-   [Intro to APIs](https://medium.com/@rwilliams_bv/apis-d389aa68104f), by Beck Williams
-   [An Illustrated Introduction to APIs](https://medium.com/epfl-extension-school/an-illustrated-introduction-to-apis-10f8000313b9), by Xavier Adam
-   [Obtaining and using access tokens for Twitter](https://cran.r-project.org/web/packages/rtweet/vignettes/auth.html)
-   [Grouped, stacked and percent stacked barplot in ggplot2](https://www.r-graph-gallery.com/48-grouped-barplot-with-ggplot2.html)

<br/>

*Note: The above readings are for when this lab used Twitter's API. Some of the information might be helpful. Please refer to class slack for Zoom recording from class.*

# Exercises

## Q1

**Learn to use an API of your choice (this can be OpenAI, which we used in class, or another model).**

```         
# Write what API you are using here.
```

## Q2

**Write code in R to authenticate with this model (hint: you might find it useful to search for R packages that help you do this, and remember: you will need some type of API key or other "credential").**

```{r}
# Your answer here


```

## Q3

**Write code to generate 25 fictional narratives less than 100 words each about a doctor and store them in a data frame.**

```{r}
# Your answer here


```

## Q4

**Write more code that uses an LLM to analyze the gender of the doctor described in each narrative.**

```{r}
# Your answer here


```

## Q5

**Report the percentage of times the LLM describes the doctor as having a "female" gender.**

```{r}
# Your answer here


```
