---
title: "Lab #3: Data Visualization"
author: "Data Science and Society (Sociology 367)"
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
In this lab we will practice using the `ggplot2` library to create visualizations of our data. **Our standard for visualizations is that each plot should have axis labels, all labels must be readable, and we should easily be able to tell what your figure is showing. Failure to do this will result in point deductions.**

Be sure to do the required readings first. While many of the problems can be solved using approaches from the lecture videos, lab videos, or required readings, you may need to do some searching on the internet to solve some of the problems. This will be a valuable skill to learn as you develop as a data scientist.

This file should be submitted in both R Markdown (.Rmd) file and knitted HTML web page (.html) files, and you should be starting with the Lab markdown file ([link to file](/assignments/Lab_2.Rmd)) where the author name has been replaced with your own. While the R Markdown file should include all of the code you used to generate your solutions, the Knitted HTML file should ONLY display the solutions to the assignment, and NOT the code you used to solve it. Imagine this is a document you will submit to a supervisor or professor - they should be able to reproduce the code/analysis if needed, but otherwise only need to see the results and write-up. This means they should be able to run the R Markdown file to directly generate the same exact html file you submitted. **Submit to your TA via direct message on Slack by the deadline indicated on the course website.**

**Required reading:** 

* [R for Data Science Chapter 5: Data transformation](https://r4ds.had.co.nz/transform.html)
* [R for Data Science Chapter 13: Relational data](https://r4ds.had.co.nz/relational-data.html)
* [`join` function documentation](https://dplyr.tidyverse.org/reference/join.html)



**Optional resources:**

- [R for Data Science (Chapters 9-13)](https://r4ds.had.co.nz/wrangle-intro.html)
- [More dplyr examples](https://rstudio.cloud/learn/primers/2)
- [More help for tidy](https://rstudio.cloud/learn/primers/4)


```{r message=FALSE, echo=FALSE}
# Load the datasets and libraries. You shouldn't need to change the URL in the `load` function.

library(tidyverse)
library(ggplot2)
library(lubridate)

load('E:\\data\\congress.RData')
```
<br/>

## Questions
<br>

**1. Describe the functionality of each of the following functions:**
```
group_py: 
summarise: 
inner_join: 
left_join: 
right_join: 
```
<br/>

**2. Compute the average age of democrat and republican (ONLY) congress members. Now do the same for M and F genders (this second part should include members of any party).**
```{r}
# your answer here
congress %>% mutate(age=2020-birthyear) %>% 
  filter(party=='Democrat' | party=='Republican') %>% 
  group_by(party) %>% 
  summarise(av_age=mean(age))

congress %>% mutate(age=2020-birthyear) %>% 
  group_by(gender) %>% 
  summarise(av_age=mean(age))


```
<br/>

**3. Create a dataframe that shows both the total number of social media accounts among democrats and republicans ONLY, as well as the within-politican averages. Which political party has more social media accounts? Which party has a higher average per-politician?**

Note: there are several ways to accomplish this. You could use `gather` again and then `group_by` and `summarise` within each individual and then within each party, or you could use mutations to get counts for each politican and then average within party.
```{r}
# your answer here
congress %>% 
  filter(party=='Democrat' | party=='Republican') %>% 
  select(twitter:youtube, party, bioguide_id) %>% 
  gather(key=sm_platform, value=sm_username, twitter:youtube) %>% 
  mutate(valid_acct=!is.na(sm_username)) %>% 
  group_by(bioguide_id) %>% 
  summarise(ct=sum(valid_acct), party=first(party)) %>% 
  group_by(party) %>% summarise(av=mean(ct), ct=sum(ct))
  
```
```
Your answers here: Which political party has more social media accounts? Which party has a higher average per-politician?
```
<br/>


**4. Use an inner join to combine the columns of the `committees` dataframe with the columns of `congress`, and then calculate the average number of committees that democrats and republicans belong to. Then calculate the averages by gender (note: this second part should include members of other parties). Do males or females tend to serve on a larger number of committees? Do republicans or democrats tend to serve on a larger number of committees?**
```{r}
# your answer here
tmp <- congress %>% 
  inner_join(committees, by=c('bioguide_id'='bioguide')) %>% 
  group_by(bioguide_id) %>% 
  summarise(n_comm=length(bioguide_id), party=first(party.x), gender=first(gender))# %>% 
  

tmp %>% 
  filter(party=='Democrat' | party=='Republican') %>% 
  group_by(party) %>% 
  summarise(av=mean(n_comm), n=length(n_comm))

tmp %>% group_by(gender) %>% 
  summarise(av=mean(n_comm), n=length(n_comm))
```
```
Your answers here: Do males or females tend to serve on a larger number of committees? Do republicans or democrats tend to serve on a larger number of committees?
```
<br/>



**5. Create a bar plot showing the number of members that belong to the 10 largest congressional committees (i.e. committees with the largest number of members). The bars should be sorted based on committee sizes.**

Note: Our standard for visualizations is that each plot should have axis labels, all labels must be readable, and we should easily be able to tell what your figure is showing. Failure to do this will result in point deductions.
```{r}
# your answer here
committees %>% group_by(committee) %>% 
  count() %>% 
  arrange(desc(n)) %>% 
  head(10) %>% 
  ggplot(aes(y=reorder(committee,n), weight=n)) + 
    geom_bar() + 
    theme_minimal() + 
    ylab('Committee') + xlab('Number of members')
```
<br/>


**6. Create a single bar plot that shows the average age of the five oldest AND youngest congressional committees (e.g. committees that have the highest and lowest average age). The bars should be sorted based on average committee ages. Which committees have the oldest and youngest average ages?**
```{r}
# your answer here
congress %>% 
  inner_join(committees, by=c('bioguide_id'='bioguide')) %>% 
  group_by(committee) %>% 
  summarise(av_age=mean(2020-birthyear)) %>% 
  arrange(desc(av_age)) %>% 
  slice(1:5, (n()-4):n()) %>% 
  ggplot(aes(y=reorder(committee,av_age), weight=av_age)) + 
    geom_bar() + 
    theme_minimal() + 
    ylab('Committee') + xlab('Average age')
```
```
Your answers here:
Which committees have the oldest and youngest averages ages?
```
<br/>




**7. Create a line graph showing the total number of politician births in each decade since the 1930's, with separate lines for senate and house members (see the `type` column). The labels on your x-axis should look like "1930's", "1940's", and so on, and your legend should show values "Senator" and "Representative" (i.e. not `rep` and `sen`).**

Note: The lines may not be continuous if there were no births in some decades.
```{r}
# your answer here
congress %>% 
  mutate(decade=birthyear %/% 10) %>% 
  group_by(decade) %>% summarise(ct=length(decade), type=first(type)) %>% 
  mutate(xticks=paste0(decade, '0\'s'), Branch=ifelse(type=='rep', 'Representative', 'Senator')) %>% 
  ggplot(aes(x=reorder(xticks,decade), y=ct, group=Branch, colour=Branch)) +
    geom_line() + 
    theme_minimal() + 
    ylab('Number of births') + xlab('Decade')

```
<br/>




**7. Use the existing `birthday` column in the `congress` dataframe to create a line graph showing the total number of politician births for each decade, with separate lines for senate and house members (see the `type` column).**

Note: the existing `birthday` column is already in POSIXct format, but you can also convert it using lubridate. Either is fine.
```{r}
# your answer here
congress %>% 
  mutate(bday=as_date(birthday)) %>% 
  
  ggplot(aes(y=reorder(committee,av_age), weight=av_age)) + 
    geom_bar() + 
    theme_minimal() + 
    ylab('Committee') + xlab('Average age')
```
```
Your answers here:
Which committees have the oldest and youngest averages ages?
```
<br/>






































**3. Change the theme of your scatterplot from Exercise 2, and apply appropriate x- and y-axis labels. Note: for theme information, see [28.6 in *R for Data Science*](https://r4ds.had.co.nz/graphics-for-communication.html#themes) or 8.3 of *Data Visualization a Practical Introduction*.**
```{r}
# your answer here
```
<br/>

**4. Now, color your scatter plot points according to whether the artists are associated with a pop genre.**
```{r}
# your answer here
```
<br/>

**5. Change the colors in the plot created for Exercise 4 to something that you think looks better. Note: for this, you may want to use a `scale_color_manual()` layer in your plot.**
```{r}
# your answer here
```
<br/>

**6. Select *and* implement an appropriate visualization to describe the relationship between `pop` and  `popularity`. Justify your choice of chart (i.e. why did you choose this chart type and not another), and describe how you would improve your chart, given more time.**
```{r}
# your answer here
```
<br/>

**7. In your own words, describe why we make data visualizations.**
```{r}
# your answer here
```
<br/>


**8. In your own words, describe at least three qualities of a good data visualization. Many of these were provided in your reading.**
```{r}
# your answer here
```
<br/>

**9. Link to a data visualization online and describe something that could make it better at conveying the central information or relationship.**
```{r}
# your answer here
```
<br/>

**10. Compare two chart types from the R Graph Gallery (`https://www.r-graph-gallery.com/index.html`). What kinds of relationships are they used to represent? What kinds of relationships would they be unable to represent? For example, do they represent continuous or categorical variable relationships? Both? Would they be useful at describing a temporal relationship, or relative proportions? These are example questions you might address as you compare and contrast. Often a chart can do many things. Simply show that you have thought deeply about how and when to use different chart types.**
```{r}
# your answer here
```
<br/>

**11. Provide three social science research questions that you would be interested in exploring for your final project. There are no right answers here. We just want to get a sense of interests.**
```{r}
# your answer here
```
