---
title: "Lab #6: Working with APIs"
author: "Data Science and Society (Sociology 367)"
date: ""
output: html_document
---


In this lab, we will practice working with application programming interfaces (APIs). Edit the Lab #6 markdown file ([link to file](https://github.com/dssoc/dssoc.github.io/raw/master/assignments/Lab_6.Rmd)) for your submission. Remember to change the "author" field above to your name. **Send the markdown file (.Rmd) to your TA via direct message on Slack.** Be sure to do the required reading!

# TODO NEED TO FILL REQUIRED READING LINKS IN

**Required reading:** [R for Data Science: Basic Programming (17-21)](https://r4ds.had.co.nz/program-intro.html).

**Optional reading:** [Intro to Network Analysis with R, by Jesse Sadle](https://www.jessesadler.com/post/network-analysis-with-r/).



### Required Datasets

We will use the following datasets for the exercises.

(1) `nominees`: [grammy_nominees_20102020.Rdata](https://github.com/dssoc/dssoc.github.io/raw/master/datasets/grammy_nominees_20102020.Rdata) is a table of Grammy nominees from 2010 to 2020 in the categories "Album of the year" (album_oty), "Record of the year" (record_oty), "Song of the year (song_oty), and "Best new artist of the year" (new_artist_oty).
(2) `winners`: [grammy_winners_20102020.Rdata](https://github.com/dssoc/dssoc.github.io/raw/master/datasets/grammy_winners_20102020.Rdata) is a table of Grammy winners for the above listed categories, from 2010 to 2020.
(3) `artist_meta`: [artist_meta.Rdata](https://github.com/dssoc/dssoc.github.io/raw/master/datasets/artist_meta.Rdata) is a table of Grammy nominee names and some of their metadata.
(4) `dnc_candidates_flw_el`: [dnc_candidates_flw_el.Rdata](https://github.com/dssoc/dssoc.github.io/raw/master/datasets/dnc_candidates_flw_el.Rdata) is an edgelist of the top five DNC cadidates for president (according to delegates won as of February 15, 2020), and the accounts they follow on Twitter.


**Load the datasets:**
```{r}
# setwd("~/Sites/dssoc_labkeys")
# library(tidyverse)
# library(spotifyr)
# library(rtweet)
# 
# load('grammy_nominees_20102020.Rdata')
# load('grammy_winners_20102020.Rdata')
# load('artist_meta.Rdata')
# load('dnc_candidates_flw_el.Rdata')
```
<br/><br/>


### API Setup

To get started, you will need to follow these setps:

1. Set up your Twitter API credentials.

(a) set up your Twitter API credentials. I


Instructions for this can be found [here](https://cran.r-project.org/web/packages/rtweet/vignettes/auth.html).





### Exercises
<br/>

**1. In your own words, describe what an application programming interface is, and what advantage it provides to data scientists/computational social scientists.**

```{r}
# Your answer here
```
<br/>

**2. Follow the steps to set up your Twitter API credentials. 

When you have completed the steps, you will enter your credentials below, but BE SURE TO REMOVE THESE BEFORE SUBMITTING YOUR ASSIGNMENT.**
```{r, eval = F}
install.packages("rjson")
library('rjson')

creds <- fromJSON(file = 'api_credentials.json')

api_key <- "<your key here>"
api_secret_key <- "<your secret here>"
access_token <- "<your token here>"
access_token_secret <- "<your token secret here>"
## authenticate via web browser
token <- create_token(
  app = "<your app name here>",
  consumer_key = api_key,
  consumer_secret = api_secret_key,
  access_token = access_token,
  access_secret = access_token_secret)
```
<br/>

**3. Using the `dnc_canidates_flw_el.Rdata` dataset, and the [`rtweet` package](https://cloud.r-project.org/web/packages/rtweet/rtweet.pdf), create a new dataset that contains additional meta information about all candidates and/or the accounts they follow. Once gathered, describe and demonstrate (e.g. in table or graph) some new insight (e.g. a summary statistic or pattern) that you gain from this new data. What more does it tell you about the candidates, the accounts they follow, or how they relate to one another?**
```{r}
# Your answer here
```
<br/>

**4. Follow [the steps](https://github.com/charlie86/spotifyr) to set up your Spotify API credentials. When you have finished, you will enter your credentials below, but BE SURE TO REMOVE THESE BEFORE SUBMITTING YOUR ASSIGNMENT.**
```{r, eval = F}
Sys.setenv(SPOTIFY_CLIENT_ID = '<your id here>')
Sys.setenv(SPOTIFY_CLIENT_SECRET = '<your secret here>')
access_token <- get_spotify_access_token()
```
<br/>

**5. Using the `get_related_artists()` command in the `spotifyR` package, create a function that takes the `spotifyID` column in the `artist_meta` dataset and returns an edge list. The edge list should include two columns: one with the Spotify ids of the Grammy-nominated artists, and one with the Spotify ids of the artists they are related to. Each row should represent a tie between one Grammy nominee, and one related artist.**
```{r}
# Your answer here
```
<br/>

**6. Now, make a network visualization of the edge list created in Exercise 5. You may do this however you'd like. Although not necessary, you may consider reformatting your data so that the network visual contains only nodes for the Grammy nominees (i.e. those Spotify ids in the `artist_meta` dataset), with edges weighted by the number of related artists they share in common. You may also consider adding artist names from the `artist_meta` dataset as node attributes in your visualization. Alternatively, you may like to gather more information from the Spotify API and add it as node or edge information in your visualization (e.g. coloring nodes by primary genre that the artist belongs to). Be creative :)**
```{r}
# Your answer here
```
<br/>


**7. Again using the `spotifyR` package, decide for yourself on some interesting data that can be added to one of the existing Grammy datasets. What new insight (e.g. a summary statistic or pattern) can you gain from this additional data? What more does it tell you about the nominees, the winners, the songs, or how any/all of these entities relate to one another?**
```{r}
# Your answer here
```
<br/>

**8. Identify another API, whether it has an associated R package or not, and describe how you might use the data available from it in a social/data scientific research project.**
```{r}
# Your answer here
```

