---
title: "Lab #5: Basics of Social Network Analysis"
author: "Data Science and Society (Sociology 367)"
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this lab we will be practicing the fundamentals of programming. 

Be sure to do the required readings first. While many of the problems can be solved using approaches from the lecture videos, lab videos, or required readings, you may need to do some searching on the internet to solve some of the problems. This will be a valuable skill to learn as you develop as a data scientist.

This file should be submitted in both R Markdown (.Rmd) file and knitted HTML web page (.html) files, and you should be starting with the Lab markdown file ([link to file](/assignments/Lab_5.Rmd)) where the author name has been replaced with your own. While the R Markdown file should include all of the code you used to generate your solutions, the Knitted HTML file should ONLY display the solutions to the assignment, and NOT the code you used to solve it. Imagine this is a document you will submit to a supervisor or professor - they should be able to reproduce the code/analysis if needed, but otherwise only need to see the results and write-up. This means they should be able to run the R Markdown file to directly generate the same exact html file you submitted. **Submit to your TA via direct message on Slack by the deadline indicated on the course website.**

**Required reading:** 

* [R4DS Chapters 17-21: Programming](https://r4ds.had.co.nz/program-intro.html)


**Optional resources:**

* [R for Data Science Chapter 16: Dates and times](https://r4ds.had.co.nz/dates-and-times.html)



```{r message=FALSE, echo=FALSE}
# Load the datasets and libraries. You shouldn't need to change the URL in the `load` function.

library(igraph)
library(ggraph)
library(tidyverse)
library(ggplot2)

load('E:\\data\\congress.RData')
```
<br/>

## Questions
<br>

**1. Describe the following concepts:**

```
nodes: 
edges: 
vertices:
links: 
directed network: 
undirected network: 
unweighted network: 
weighted network: 
bipartite network: 
```
<br/>

**2. What kinds of questions would you be able to answer with networks that you might not be able to answer using analysis techniques we have covered so far?**
```
# Yor answer here or in text.
```
<br/>

**3. Describe the data in the `candidates_flw_am` adjacency matrix and the `candidates_flw_el` edgelist. How do the two networks differ? Do they represent directed or undirected edges? Do the edges have weights?**
```
# Yor answer here or in text.
```
<br/>

**4. Describe three different metrics (e.g. betweenness centrality, degree, etc) that can be used to summarize networks or specific nodes/vertices within a network.**
```{r}
# Your answer here
```
<br/>


**5. Using the `candidates_flw_am` *adjacency matrix*, create a graph (network) object. Make sure to include information about the edge weights (i.e. the number of accounts shared in common). Hints: (1) you may need to use the `as.matrix` function to the dataframe before converting to igraph if you are having problems with data types; (2) you may need to look through `igraph`'s official documentation (see link under "optional reading" above) to decide which function will convert adjacency matrices to igraph objects.**
```{r}
# Yor answer here or in text.
```
<br/>


**6. Using the graph object that you created, indicate which candidate has the highest betweenness centrality, taking into account edge weight. If you had to write this finding in a research paper, how would you describe its meaning to a social science audience?**
```{r}
# Yor answer here or in text.
```
<br/>


**7. Make a visual of your graph object using `ggraph`, with edges scaled by weight (i.e. the number of accounts that the two candidates mutually follow). If you decide to make this graph in Gephi, feel free to submit a separate image of your final visual in addition to your .Rmd file.**
```{r}
# Yor answer here or in text.
```
<br/>

**8. Now we turn our attention to two-mode networks. Create an igraph object from the `candidates_flw_el` edgelist, and calculate the degree for each candidate.**
```{r}
# Yor answer here or in text.
```
<br/>


**9. In last week's lab excerices, you were asked to indicate possible research topics that you would be interested in pursuing for your final project. Choose one of your topics and expand on it in more detail (i.e. at least one paragraph).**
```{r}
# your answer here
```



