---
title: "Lab #1"
author: "Data Science and Society (Sociology 367)"
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<br>

Please complete as many of the questions below as you can. If you are not sure how to write code to answer them, write comments in each code chunk explaining where you got stuck so that we can give you help you improve and/or give you partial credit. To submit your answer, you'll need to edit this markdown document (you can copy and paste the text form [this link](/assignments/Lab_1.Rmd) into a new `.Rmd` file or write a new one from scratch. If you copy and paste, Rremember to change the "author" field above to your name. After you've edited it, save it and then send it to your TA via direct message on Slack. To learn how to load and edit markdown files in R, I've created [this video tutorial](https://youtu.be/DkqFH0C_CPQ) for you. To complete these exercises, you'll need to download the `Coronavirus_Tweets.Rdata` dataset from [this link.](/datasets/Coronavirus_Tweets.Rdata)


This file should be submitted in both R Markdown (.Rmd) file and knitted HTML web page (.html) files, and you should be starting with the Lab markdown file ([link to file](/assignments/Lab_10.Rmd)) where the author name has been replaced with your own. While the R Markdown file should include all of the code you used to generate your solutions, the Knitted HTML file should ONLY display the solutions to the assignment, and NOT the code you used to solve it. Imagine this is a document you will submit to a supervisor or professor - they should be able to reproduce the code/analysis if needed, but otherwise only need to see the results and write-up. This means they should be able to run the R Markdown file to directly generate the same exact html file you submitted. **Submit to your TA via direct message on Slack by the deadline indicated on the course website.**



**Required reading:** 

* [R for Data Science (Chapters 4, 5, 10)](https://r4ds.had.co.nz/workflow-basics.html)


## Dataset Description
https://github.com/unitedstates/congress-legislators



**Load the datasets and libraries. You shouldn't need to change the URL in the `load` function**
```{r message=FALSE, echo=FALSE}
library(tidyverse)
library(readxl)

congress_info <- read_excel('E:\\data\\legislators-current(1).xlsx')

# THIS SHOULD WORK AS-IS
#load(url('https://dssoc.github.io/datasets/Coronavirus_Tweets.Rdata'))
#load(url('https://dssoc.github.io/datasets/congress.RData'))
load('E:\\data\\congress.RData')

#congress_info <- read.csv(url('https://theunitedstates.io/congress-legislators/legislators-historical.csv'), stringsAsFactors = FALSE)

#congress_info <- read.csv(url('https://theunitedstates.io/congress-legislators/legislators-historical.csv'), stringsAsFactors = FALSE)

```
<br/>



<br>
**1. Load the "Coronavirus_Tweets.Rdata" dataset. Write code that calculates the mean value of the variable "retweet_count" in the code chunk below**

```{r message=FALSE}
#write your code here
congress %>% mutate(birthyear=as.numeric(substr(birthday, 1, 4))) %>% select(birthyear)
```
<br>
**2. Why does the code below create an error message? Re-write the code so that it does not create an error message.**

```{r, eval=FALSE}
My_Variable<-1
My_Var1able
```
<br>
**3. Extract the third element of the vector.**

```{r}
Vector_Saurus_Rex<-c("R","is","great")
```
<br>
**4. Write code that identifies the largest number in this vector (hint: you'll have to research some new functions in r to do this- try Google or one of the tutorials linked on the syllabus).**

```{r}
Large_Vector<-c(25555,342343,123123123,4234234,53243234,54324234,5421111,12312312,111231,
                     1231231,12312312,12312312,123123,898972,789872,2343,23423423,2343221,23423,
                     14444,44324222,2342341,124231111,22233345,1111233333,1231231,1231231)
```
<br>
**5. Calculate the square root of 2456.**
```{r}
# write your code here
```
<br>
**6. Which Tweet in the Coronavirus Tweets.Rdata dataset received the most "likes" (or "favorites"? Write code that identifies the tweet and displays the text of the tweet.**
```{r}
# write your code here
```
<br>
<br>
<br>
