---
title: "Lab #0: Example R Markdown File"
author: "Data Science and Society (Sociology 367)"
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**[Lab #0 Markdown File](/assignments/Lab_0_example_markdown.Rmd)**

You can use this example lab file to make sure you can Knit .Rmd files in RStudio. This lab will not be graded, so don't worry about submitting it. Feel free to ask the TA if you are unsure about the submission format.

**Each week's lab should be submitted to your TA via Slack direct message by the deadline indicated on the course website.** You should start with the markdown file linked above, and you will submit two files for each lab: an [R Markdown (.Rmd) file](https://rmarkdown.rstudio.com/authoring_quick_tour.html) containing solution code and written responses that answer each question, and an HTML file (.html) generated by knitting that file. The TA should be able to knit your R Markdown file to reproduce your html file exactly without any additional steps, and the knitted HTML file should display your code and ONLY the output needed to answer the question. The output from your solution code should be easy to read and you will lose points if your knitted html file includes extraneous output that makes your solution harder to read.

**Required reading:** 

Any required readings for the labss will be listed here.

**Optional resources:**

Optional readings for the labs will be listed here.


## Setup

Every lab will have some code that loads several libraries and the dataset you should use for the lab. If you need to load other libraries to generate your solutions, you can add that code here - I just wrote these to get you started.

```{r message=FALSE, echo=FALSE}
# Load the datasets and libraries. You shouldn't need to change the URL in the `load` function.

library(tidyverse)

# this will download the dataset from the course website as a dataframe called `congress`.
load(url('https://dssoc.github.io/datasets/congress.RData'))
```
<br/>


## Example Questions

<br>

**1. List the names of each column in the `congress` dataset we loaded above.**
```
colnames(congress)
```
<br>
