---
title: "Lab #3: Data Visualization"
author: "Data Science and Society (Sociology 367)"
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this lab, we will practice visualizing data. Edit the Lab #3 markdown file ([link to file](/assignments/Lab_3.Rmd)) for your submission. Remember to change the "author" field above to your name. Send the markdown file (.Rmd) to your TA via direct message on Slack.

The exercises in this lab correspond with  *R for Data Science* (Wickham), [chapters 3 and 28](https://r4ds.had.co.nz/data-visualisation.html), and *Data Visualization a Practical Introduction* (Healy), [chapter 1](http://socviz.co/lookatdata.html).

**Optional Resources:**

- [RStudio Cloud intro](https://rstudio.cloud/learn/primers/3)
- [Data Visualization: A Practical Introduction](http://socviz.co/), Kieran Healy. (Princeton: Princeton University Press, 2019)
- [Fundamentals of Data Visualization](https://serialmentor.com/dataviz/)
- [The R Graph Gallery](https://www.r-graph-gallery.com/)
- [Sometimes the y-axis doesn't start at zero, and it's fine.](https://flowingdata.com/2015/11/23/sometimes-the-y-axis-doesnt-start-at-zero-and-its-fine/)
- [What would feminist data visualization look like?](https://civic.mit.edu/2015/12/01/feminist-data-visualization/), MIT Civic Media Lab by kanarinka. 


### Required Dataset

The data that you will be using for this lab can be downloaded from the course website: 

- `artist_meta`: [artist_meta.Rdata](/datasets/artist_meta.Rdata) is a table of Grammy nominee names and some of their metadata.


**Load the datasets:**
```{r message=FALSE}
#library(tidyverse)
#load('artist_meta.Rdata')
library(tidyverse)
load('../datasets/artist_meta.Rdata')
print(artist_meta)
```
<br/><br/>

### Exercises
<br/>

**1. Using the `artist_meta` table, create a new factor variable (`is_pop`) that indicates whether the artist is associated with a pop genre. The labels for this factor variable should be 'pop' and 'not pop'. Note that any genre that includes the word "pop" should be considered as pop. Hint: consider using the `grepl` and `ifelse` functions together, and a type conversion to get a factor variable. See "R for data science" Chapter 15 for more information about factor variables.**

**Also notice one additional issue with the data that might cause issues later on: the `popularity` column is an integer list of size one. Consider using the `mutate` function with the `is.numeric` function to coerce the data into the type you need (see section 20.4.1 of *R for Data Science* to learn more).**
```{r}
# your answer here
artist_meta <- artist_meta %>% 
  mutate(popularity=as.numeric(popularity)) %>% 
  mutate(is_pop = as.factor(ifelse(grepl('pop', as.character(genre), fixed=TRUE), 'pop', 'not pop')))
print(artist_meta)
```
<br/>

**2. Using the `artist_meta` table, make a scatterplot of `followers` vs `popularity`.**
```{r}
# your answer here
artist_meta %>% ggplot(aes(followers, y=popularity)) + 
  geom_point()
```
<br/>

**3. Change the theme of your scatterplot from Exercise 2, and apply appropriate x- and y-axis labels. Note: for theme information, see [28.6 in *R for Data Science*](https://r4ds.had.co.nz/graphics-for-communication.html#themes) or 8.3 of *Data Visualization a Practical Introduction*.**
```{r}
# your answer here
artist_meta %>% ggplot(aes(followers, y=popularity)) + 
  geom_point() + 
  labs(x= 'Number of followers', y='Popularity score')
```
<br/>

**4. Now, color your scatter plot points according to whether the artists are associated with a pop genre.**
```{r}
# your answer here
artist_meta %>% ggplot(aes(followers, y=popularity)) + 
  geom_point(aes(colour = is_pop)) + 
  labs(x= 'Number of followers', y='Popularity score', colour='Is a pop genre?')
```
<br/>

**5. Change the colors in the plot created for Exercise 4 to something that you think looks better. Note: for this, you may want to use a `scale_color_manual()` layer in your plot.**
```{r}
# your answer here
artist_meta %>% ggplot(aes(followers, y=popularity)) + 
  geom_point(aes(colour = is_pop)) + 
  labs(x= 'Number of followers', y='Popularity score', colour='Is a pop genre?') + 
  scale_color_manual(values=c("#999999", "#E69F00"))
```
<br/>

**6. Select *and* implement an appropriate visualization to describe the relationship between `pop` and  `popularity`. Justify your choice of chart (i.e. why did you choose this chart type and not another), and describe how you would improve your chart, given more time.**
```{r}
# your answer here
artist_meta %>% ggplot(aes(followers, y=popularity)) + 
  geom_point(aes(colour = is_pop)) + 
  labs(x= 'Number of followers', y='Popularity score', colour='Is a pop genre?') + 
  scale_color_manual(values=c("#999999", "#E69F00")) + 
  geom_smooth(method='glm', formula = y ~ x + x**2)
```
<br/>

**7. In your own words, describe why we make data visualizations.**
```{r}
# your answer here
```
<br/>


**8. In your own words, describe at least three qualities of a good data visualization. Many of these were provided in your reading.**
```{r}
# your answer here
```
<br/>

**9. Link to a data visualization online and describe something that could make it better at conveying the central information or relationship.**
```{r}
# your answer here
```
<br/>

**10. Compare two chart types from the R Graph Gallery (`https://www.r-graph-gallery.com/index.html`). What kinds of relationships are they used to represent? What kinds of relationships would they be unable to represent? For example, do they represent continuous or categorical variable relationships? Both? Would they be useful at describing a temporal relationship, or relative proportions? These are example questions you might address as you compare and contrast. Often a chart can do many things. Simply show that you have thought deeply about how and when to use different chart types.**
```{r}
# your answer here
```
<br/>

**11. Provide three social science research questions that you would be interested in exploring for your final project. There are no right answers here. We just want to get a sense of interests.**
```{r}
# your answer here
```
