---
title: "Lab #7: Modeling with Statistics"
author: "Data Science and Society (Sociology 367)"
date: ""
output: html_document
---

**[Lab #7 Markdown File](/assignments/Lab_7.Rmd)**

**[Lab Instructions](https://dssoc.github.io/assignments/lab_instructions.html)**

In this lab, we will practice making and visualizing statistical models of data using R.

See the "Instructions" section of the [Introduction to Lab Assignments](https://dssoc.github.io/assignments/lab_instructions.html) page for more information about the labs. That page also gives descriptions for the datasets we will be using.

**Required reading:** 

* R for Data Science: [Modeling (Chapters 23-25)](https://r4ds.had.co.nz/model-intro.html)
* [Quick Guide: Interpreting Simple Linear Model Output in R](https://feliperego.github.io/blog/2015/10/23/Interpreting-Model-Output-In-R)


**Optional reading:** 



```{r message=FALSE, echo=FALSE}
# Load the datasets and libraries. You shouldn't need to change the URL in the `load` function.

library(tidyverse)
library(ggplot2)
library(modelr) # this one is new!

load(url('https://dssoc.github.io/datasets/congress.RData'))
load(url('https://dssoc.github.io/datasets/committees.RData'))
```
<br/>


## Questions
<br>

**1. In your own words, describe what statistical modeling means. When is it used? What does it allow data scientists to do?**

```
# Your answer here
```
<br/>

**2. Create the following new variables for each congress member: (a) age of each congress member, (b) the number of subcommittees that each congress member is on, and (c) the percentage of instances where congress members hold a title in the subcommittees they belong to (i.e. when the `title` entry in the `committee_membership` dataframe is not empty). You will want to save these new variables for future problems. Then use the `summary` function to create summary statistics for (ONLY) these new variables.**
```{r}
# Your answer here
```
<br/>

**3. Create a scatter plot with a linear trendline (must be a straight line) to predict the number of full committees that congress members belong to from `age`. Describe the relationship. What do each of these (the scatter points and the linear trendline) show that you cannot see from the other? **

Note: usually we see the dependent variable (number of committees in this case) on the y-axis and the independent variable (age in this case) on the x-axis.
```{r}
# Your answer here
```
```
written explanation here
```
<br/>

**4. Create a bar graph showing the average number of full committees that congress members belong to by gender (i.e. a bar for M and a bar for F) with error bars. What can you see from this visualization? Does there appear to be a significant difference?**

Hint: you may want to see `geom_errorbar`.

Hint: error bars are usually calculated by taking the average plus and minus the standard deviations.
```{r}
# Your answer here
```
```
written explanation here
```
<br/>



**5. Following section [24.2.2](https://r4ds.had.co.nz/model-building.html#a-more-complicated-model) of the R4DS required reading, construct a model using `lm` or `glm` to predict the percentage of time that a member holds a title in the committees they belong to from age, gender, and political party. Keep this model for future problems. Based on the model `summary`, which variables might be related to holding a title? Try removing and adding different variables. Does changing any of the used variables change your original interpretation?**

HINT: see required readings for help interpreting regression models.
```{r}
# Your answer here
```
```
written explanation here
```
<br/>

**6. Following that same R4DS section, use `data_grid` to create a dataframe where `age` and `gender` vary and political party is held constant, and then use `add_predictions` to make a new column for predictions based on our model. Using that dataframe, make a plot showing model predictions from age with separate lines for the two genders. How do you interpret this plot?**

HINT: check the arguments for `data_grid` to specify which variables to vary and which to hold constant. You'll likely want to use the `.model` argument.
```{r}
# Your answer here
```
```
Your answer here
```
<br/>

**7. How could you use statistical modeling to answer the hypothesis you provided in the last assignment? What inferences could you make?**
```
# Your answer here
```
<br/>


**8. Describe one or two existing datasets that you would like to use for the project you've been developing last week. Will you be able to download the data from somewhere, or can you use an API? Will you be making statistical models, analyzing networks, doing text analysis, or creating visualizations? See the ["Final Project " section in the course description page on the website](https://dssoc.github.io/course/#formal-requirements).**
```
# Your answer here
```
<br/>