---
title: "Lab #1: R Basics"
author: "Data Science and Society (Sociology 367)"
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**[Lab #1 Markdown File](/assignments/Lab_1.Rmd)**

**[Lab Instructions](https://dssoc.github.io/assignments/lab_instructions.html)**

In this lab we will be learning the basics of using R and working with real data.

See the "Instructions" section of the [Lab Instructions](https://dssoc.github.io/assignments/lab_instructions.html) page for more information about the labs. That page also gives descriptions for the datasets we will be using.


**Required reading:** 

* [R for Data Science Chapter 4: Workflow basics](https://r4ds.had.co.nz/workflow-basics.html)
* [R4DS Chapters 26-30: Communicating your analysis](https://r4ds.had.co.nz/communicate-intro.html) (Except Chapter 28 - we will cover visualization later.)


**Optional resources:**

* [RDocumentation.org]([documentation website](https://www.rdocumentation.org/))



```{r message=FALSE, echo=FALSE}
# Load the datasets and libraries. You shouldn't need to change the URL in the `load` function.

library(tidyverse)
library(readxl)

# this will download the dataset from the course website
load(url('https://dssoc.github.io/datasets/congress.RData'))
```
<br/>


## Questions

<br>



**1. Use the [documentation website](https://www.rdocumentation.org/) or the helper functions in R (`?function_name`) to look up the `mean` function, and describe each of its arguments and what they do in your own words. Write your answer in the code block below as a comment.**

NOTE: You can write comments in a code block by including a "#" character at the beginning of the line.
```{r}
# Write your answer here.
```
<br>


**2. Extract the third element of the vector using numerical indexing.**

NOTE: the first element index starts with 1 in R.
```{r}
my_vector <- c("R","is","great")
# your code here
```
<br>

**3. This question has several parts: (a) Use R code to identify the data type of `some_vector`. (b) What is the largest number in this vector? (c) How about the mean value? (d) Does this vector include the number `54324222`? Your results should be generated from the R code - please do not write the numbers into R manually.**

HINT: you'll have to research some new functions in r to do this- try Google or one of the tutorials linked on the syllabus.

HINT: check out the `%in%` operator.
```{r}
some_vector <- c(25555,342343,123123123,4234234,53243234,54324234,5421111,12312312,111231,
                     1231231,12312312,12312312,123123,898972,789872,2343,23423423,2343221,23423,
                     14444,44324222,2342341,124231111,22233345,1111233333,1231231,1231231)
# your code here
#a
class(some_vector)
#b
max(some_vector)
#c
mean(some_vector)
#d
54324222 %in% some_vector
```
<br>


**4. Use the `congress` dataframe for the remainder of questions in this assignment. (a) How many rows and columns does the `congress` dataframe have? (b) Use a function to show its data type. You must use R code to generate these values.**

```{r}
# your code here
nrow(congress)
ncol(congress)
```
<br>


**5. Using the browse features of R or RStudio, write the ages of Sherrod Brown (a member of congress) and Dianne Feinstein.**

```
Your written response here.
Sherrod Brown: 
Dianne Feinstein: 
```
<br>


**6. What is the average birthyear of congress members? What is the data type of the `birthyear` column?**

```{r}
# your code here
mean(congress$birthyear)
class(congress$birthyear)
```
<br>

**7. What are your learning goals for this semester? Not a trick question ;)**
```
your written response here
```
