---
title: "Lab Instructions and Data Descriptions"
author: "Data Science and Society (Sociology 367)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

In this document I'll first give detailed instructions for the labs and then give descriptions for each dataset we will be using this semester.


# Instructions

**Each week's lab should be submitted to your TA via Slack direct message by the deadline indicated on the course website.** You should start with the markdown file linked at the top of the lab, and you will submit two files for each lab: an [R Markdown (.Rmd) file](https://rmarkdown.rstudio.com/authoring_quick_tour.html) containing solution code and written responses (if required) that answer each question, and an HTML file (.html) generated by knitting that file. The TA should be able to knit your R Markdown file to reproduce your html file exactly without any additional steps, and the knitted HTML file should display your code and ONLY the output needed to answer the question. The output from your solution code should be easy to read and you will lose points if your knitted html file includes extraneous output that makes your solution harder to read.

**Be sure to do the required readings first.** While many of the problems can be solved using approaches from the lecture videos, lab videos, or required readings, you may need to do some searching on the internet to solve some of the problems. This will be a valuable skill to learn as you develop your data science skillset.

# Lab File List

This is a list of all Labs and their associated markdown files:

0. [Example R Markdown File (ungraded)](/assignments/Lab_0.html) ([markdown](/assignments/Lab_0.Rmd))

<!--
1. [R Basics](/assignments/Lab_1.html) ([markdown](/assignments/Lab_1.Rmd))
2. [Data Visualization](/assignments/Lab_2.html) ([markdown](/assignments/Lab_2.Rmd))
3. [Data Wrangling](/assignments/Lab_3.html) ([markdown](/assignments/Lab_3.Rmd))
4. [Programming Basics](/assignments/Lab_4.html) ([markdown](/assignments/Lab_4.Rmd))
5. [Coding Social Networks](/assignments/Lab_5.html) ([markdown](/assignments/Lab_5.Rmd))
6. [Working with APIs](/assignments/Lab_6.html) ([markdown](/assignments/Lab_6.Rmd))
7. [Modeling (A Brief Introduction)](/assignments/Lab_7.html) ([markdown](/assignments/Lab_7.Rmd))
8. [Introduction to text analysis](/assignments/Lab_8.html) ([markdown](/assignments/Lab_8.Rmd))
9. [Word counts and Dictionaries](/assignments/Lab_9.html) ([markdown](/assignments/Lab_9.Rmd))
10. [Topic Modeling](/assignments/Lab_10.html) ([markdown](/assignments/Lab_10.Rmd))
-->

# Dataset Description

Now I will describe the datasets we will use for the class. We will use a total of three different [RData files](https://bookdown.org/ndphillips/YaRrr/rdata-files.html):


+ [`congress.RData`](https://dssoc.github.io/datasets/congress.RData): contains basic information about each member of congress.
+ [`senator_tweets.RData`](https://dssoc.github.io/datasets/senator_tweets.RData): a small sample of Tweets from accounts associated with each senator (a subset of all congress members) in the `congress` dataset.
+ [`congress_wiki.RData`](https://dssoc.github.io/datasets/congress_wiki.RData): text from wikipedia pages of each member of congress in the `congress` dataset.



## The `congress.RData` Dataset

[Download congress.RData](https://dssoc.github.io/datasets/congress.RData)

This dataset contains information about each member of congress that was in-office as of January 11, 2021 and their committee memberships. It was retrieved from the [congress-legislators repository](https://github.com/unitedstates/congress-legislators), a database "maintained through a combination of manual edits by volunteers (from GovTrack, ProPublica, MapLight, FiveThirtyEight, and others) and automated imports from a variety of sources." From this source we use the `legislators-current` and 


You can see the January 11, 2021 

that was collected by volunters from sources []()

legislators-current

https://github.com/unitedstates/congress-legislators/tree/5b25a97ca2d93f51a4a06dd87cdb16d4d3052c10


You will use this line of code to download the Rdata file from the course website and open it directly in RStudio. In theory, you could also download the file and point the `load()` function to that file on your computer, but please use this line so your code is easily reproducible on any computer.

```{r}
load(url('https://dssoc.github.io/datasets/congress.RData'))
```

This dataset consists of three different variables:

+ `congress`: basic information like name, birthday, state of representation, political party, and social media account handles for each of the 536 members of the Xth congress. 
+ `committees`: information about committee membership for each member of congress.
+ `congress_ids`: detailed lookup information for each member of congress.

### `congress` Dataframe

Each row in this dataframe corresponds to a member of congress

```{r}
congress %>% head()
```

### `committees` Dataframe

```{r}
committees %>% head()
```

### `congress_ids` Dataframe

```{r}
congress_ids %>% head()
```

## The `senator_tweets.RData` Dataset

[Download senator_tweets.RData](https://dssoc.github.io/datasets/senator_tweets.RData)


## The `congress_wiki.RData` Dataset

[Download congress_wiki.RData](https://dssoc.github.io/datasets/congress_wiki.RData)



