---
title: "Lab #3: Data Visualization"
author: "Data Science and Society (Sociology 367)"
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**[Lab #3 Markdown File](/assignments/Lab_3.Rmd)**

**[Lab Instructions](https://dssoc.github.io/assignments/lab_instructions.html)**

In this lab we will practice using the `ggplot2` library to create visualizations of our data. **Our standard for visualizations is that each plot should have axis labels, all labels must be readable to someone unfamiliar with your data (e.g. Female and Male instead of `F` and `M`, Senator and Representative instead of `rep` and `sen`), and we should easily be able to tell what your figure is showing. Failure to do this will result in point deductions.**

See the "Instructions" section of the [Introduction to Lab Assignments](https://dssoc.github.io/assignments/lab_instructions.html) page for more information about the labs. That page also gives descriptions for the datasets we will be using.

**Required reading:** 

* [R for Data Science Chapter 13: Relational data](https://r4ds.had.co.nz/relational-data.html)
* [R for Data Science Chapter 3: Data visualization](https://r4ds.had.co.nz/data-visualisation.html)
* [R for Data Science Chapter 28: Graphics for communication](https://r4ds.had.co.nz/graphics-for-communication.html)


**Optional resources:**

* [R for Data Science Chapter 16: Dates and times](https://r4ds.had.co.nz/dates-and-times.html)
* [Grouped, stacked and percent stacked barplot in ggplot2](https://www.r-graph-gallery.com/48-grouped-barplot-with-ggplot2.html)


```{r message=FALSE, echo=FALSE}
# Load the datasets and libraries. You shouldn't need to change the URL in the `load` function.

library(tidyverse)
library(ggplot2)
library(lubridate)

load(url('https://dssoc.github.io/datasets/congress.RData'))
load(url('https://dssoc.github.io/datasets/committees.RData'))
```
<br/>


## Example Questions
<br>


**ex1. make a bar chart showing the number of male and female members of congress in our dataset.**
```{r}
congress %>% 
  ggplot(aes(x=gender)) +
    geom_bar()
```
<br/>


**ex2. make a bar chart showing the proportion of female senators in each pol.**
```{r}
congress %>% 
  group_by(party) %>% 
  summarize(proportion_gender=mean(gender=='F')) %>% 
  ggplot(aes(x=party, y=proportion_gender)) +
    geom_bar(stat='identity', position='dodge')
```
<br/>






## Questions
<br>

**1. Describe the functionality of each of the following functions:**
```
group_by: 
summarise: 
inner_join: 
left_join: 
ggplot: 
```
<br/>

**2. Create a bar plot to show the average ages of congress members from each political party. Now do the same for M and F genders.**
```{r}
# your answer here
```
<br/>


**3. Create a line graph showing the total number of congress member births in each decade since the 1930's, with separate lines for senate and house members (see the `type` column). The labels on your x-axis should look like "1930's", "1940's", and so on, and your legend should show names "Senator" and "Representative" (i.e. not `rep` and `sen`).**

Note: The plotted lines may not show up in decades where there were no births - that is okay.
```{r}
# your answer here
```
<br/>

**4. Create a bar chart showing the average ages of Senators and Representatives separately by weekday. The plot should make it easy to compare Senators and Representatives within each weekday. The x-labels should be weekday names and appear in chronological order.**

NOTE: For convenience, I have already parsed the `birthdate` column into a `date` type, so you'll just need to extract the day of week using `lubridate` methods.

HINT: the final plot should have 14 bars: 7 weekdays by 2 types of congress members (Senators and Representatives).

HINT: see the Optional Readings for more information about grouped bar charts using `ggplot2`.
```{r}
# your answer here
```
<br/>


**5. Use an inner join to combine the columns of the `congress` dataframe with the columns of `congress_contacts`, and show the average number of valid Twitter, Facebook, and YouTube social media accounts among congress members of each political party as a bar chart. Do the same for F and M genders.**

HINT: you will want to join the dataframes based on a column that is common to both datasets.

HINT: there are several ways to accomplish the counting of social media accounts. You could use `gather` again and then `group_by` and `summarise` within politician and then within party, or you could use `mutate` to get counts for each politican and then average by party. Any other approach is also fine.
```{r}
# your answer here
```
<br/>


**6. The remainder of problems in this homework will involve the use of the `committees.RData` dataset. You should refer to the Lab Instructions page for more information about this dataset. Create a plot showing the average number of subcommittees that F and M genders belong to. Then create a plot showing the average number of full committees that congress members belong to by party.**

HINT: as described in the Lab Instructions page, you will need to perform a join between the `subcommittees` and `committee_memberships` dataframes to get ONLY subcommittee (and not committee) memberships, and a join between the `committees` and `committee_memberships` to get ONLY full committees. You may copy-paste the code from the Lab Instructions page if that would be helpful.

```{r}
# your answer here
```


**7. Create a bar plot showing the number of members that belong to the 5 largest full congressional committees (i.e. full committees with the largest number of members). The bars should be sorted based on committee sizes.**

NOTE: read the Lab Instructions page for more information about the standards for visualizations in this course. The full committee names should appear somewhere on the plot - please do not provide `thomas_id`s only (you may include full committee names in the legend though). 

HINT: See the Lab Instructions page for more info about filtering for full committees only (excluding subcommittees).
```{r}
# your answer here
```
<br/>


**8. Create a single bar plot that shows the average age of the full committees with the 5 highest and lowest average ages. The bars should be sorted based on average committee ages.**

This means you will need to join three dataframes: `committee_memberships` for membership information, `committees` to separate full committees from subcommittees and get committee names, and `congress` to get age information.

```{r}
# your answer here
```
<br/>

**9. Using the topics you described last week or a new topic you have been thinking about, describe two social science questions that you would be interested in exploring for your final project. Do you think these questions might be answerable using real data?**
```{r}
# your answer here
```











